### 2019-05-29 paper.tex
Dr. Antonio edits.

%\documentclass[twocolumn,letterpaper]{article}
\documentclass[10pt]{csce}

\usepackage[hmargin=.75in,vmargin=1in]{geometry}
\usepackage[american]{babel}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{caption}
%\usepackage{flushend}

\usepackage{gensymb,hyperref,graphicx,paralist,amsmath,multirow}
\usepackage[table]{xcolor}
\pagenumbering{gobble}

\title{\bf Enhancing Scheduling Robustness with Partial Task Completion
Feedback and Resource Requirement Biasing}

%%%% Make sure the author names are boldface.
\author{
{\bfseries Nicolas Grounds and John K. Antonio}\\
School of Computer Science, University of Oklahoma, Norman, Oklahoma, United States of America\\
Research Paper
}

\begin{document}

\maketitle                        %%%% To set Title and Author names.

\begin{abstract}
Performance and robustness of dynamic scheduling algorithms are evaluated in
the presence of errors in the tasks' resource requirements.  Previous work
found that incorporating task completion events from the actual distributed
system into the algorithms' model of the system was crucial for achieving
robustness.  In the present paper, various degrees of feedback, rather than
simply all-or-none, are evaluated using the same simulated studies as in
previous work and a proposed strategy for biasing model tasks' resource
requirement information is proposed in order to counteract the most egregious
effects of model error on performance.
\end{abstract}

\vspace{1em}
\noindent\textbf{Keywords:}
 {\small distributed system, scheduling, performance, robustness, biasing}

\section{Introduction and Background}
\label{sec:Intro}

Scheduling computational tasks to machines so as to improve specified metrics
of performance has been the topic of a plethora of good work produced over the
past several decades \cite{taxonomy}. The underlying assumptions and objectives
of this body of work varies along several dimensions. First, some work assumes
all tasks are independent whereas other work, as in this paper, allows for
tasks to have dependency or precedence relationships with other tasks (for
which interrelated tasks are typically represented in a directed acyclic graph,
or DAG). Additionally, there are static formulations to scheduling in which a
desired schedule is detemined offline based on assumed knowledge related to the
machines' available resources and, correspondingly, the resource requirements
of the computational tasks.

This paper addresses dynamic scheduling in which the schedule for
tasks is determined online in real-time with the execution of those tasks
performed on a distributed system. Unlike static scheduling, dynamic scheduling
does not require upfront knowledge of the arrival of future tasks into the
system for scheduling. Algorithms for dynamic scheduling make use of knowledge
about the tasks which are ready for scheduling and their resource requirements
such as CPU and memory load as well as the resource capacities of the machines
in the distributed system.

Another dimension in the taxonomy of scheduling algorithms deals with whether
tasks are executed `one at a time' on the machine to which they are assigned.
In the present work, multiple tasks may be executed concurrently on a single
machine. This adds complexity in modeling the machines' performance because
the machines' resources must be shared across multiple tasks assigned to the
machine.

Some algorithms for dynamic scheduling are based on heuristics for selecting
which tasks to prioritize and determining when to begin their execution and on
which machine. Other algorithms attempt to optimize scheduling decisions with
respect to a desired outcome based on a user-defined objective.  Both types of
algorithms are generally measured and compared to one another against such
objectives as minimizing makespan (time required for completed execution
of all tasks) \cite{stochastic}, or, as in this paper, the degree to which all
tasks of a DAG are completed by a DAG-associated deadline.

Scheduling algorithms may orthogonally be evaluated based on their robustness,
for example, how well the same objective is achieved when information provided
to the scheduling algorithm contains errors such as inaccuracies in resource
requirements of the tasks.  In previous work \cite{pdpta18} four dynamic
scheduling algorithms' robustness to error with respect to performance against
an objective of completing DAGs before their deadline was presented, showing
how some algorithms were not robust to even the smallest amount of error. The
use of task completion event feedback from the actual distributed system back
into the modeled system used by the scheduling algorithms was found to
substantially improve robustness of all four scheduling algorithms even with
large amounts of error. In the present paper we generalize this approach by
investigating the utility of using only partial feedback of task completion
events.

The remainder of the paper is organized in the following manner.  Section
\ref{sec:Framework} describes the problem domain and the simulation software's
modeling of a distributed system in which errors in task requirements may
be present.  Section \ref{sec:Biasing} presents a requirements biasing approach
to counteracting model error to prevent scheduling algorithms from
over-committing system resources, i.e., executing too many tasks concurrently.
Section \ref{sec:Results} presents results of simulated case studies.
Finally, Section \ref{sec:Summary} summarizes the findings from these
simulations and presents the conclusions of our work.

\section{Problem Domain and \\
Simulation Environment}
\label{sec:Framework}

The present work is an extension to previous work \cite{pdpta18} in which a
model-based approach to dynamically scheduling tasks from DAGs (called
workflows) was introduced.  Four scheduling algorithms were tested in a
simulation environment \cite{soasim} given a 24-hour simulated period
of arriving workflows ranging in size from 5 tasks up to 800.  Each workflow
has a known, predetermined deadline and scheduling algorithms were evaluated
in \cite{pdpta18} based both on the number of workflows completed before their
deadline and the distribution of workflow counts completed at various
normalized proportions past their deadline (e.g., workflows up to 100\% late
relative to the amount of time between their arrival and deadline).

Scheduling algorithms are used to determine \textemdash\ from a given queue of tasks
ready to begin executing (i.e., tasks of arrived workflows that have no
precedence constraints or whose precedent tasks have all completed)
\textemdash\ both when the task should begin executing and on which machine of
the platform. Unlike some scheduling research, tasks are permitted to executed
concurrently with other tasks on the same machine, which increases machines'
overall load on resources such as CPU and memory and thereby slows the rate of
work on each executing task.  Original work in \cite{pdpta09} details this
non-linear degradation of rate of work (efficiency) due to concurrent task
execution.

Figure \ref{fig:platform} illustrates the various components and general
flow of information within the model-based approach to executing and evaluating
scheduling algorithms.  The modeled tasks' requirements (CPU load,
required number of CPU cycles, and memory load) may contain errors relative to
the true values; these errors then cause the model platform to
diverge from the actual platform in terms of which tasks are completed and
which are still executing.

\begin{figure}
	\begin{center}
		\includegraphics[angle=-90,width=0.5\textwidth]{figures/PlatformDiagram.pdf}
	\end{center}
	\caption{Block diagram illustrating model-based framework from
		\cite{pdpta18}.}
	\label{fig:platform}
\end{figure}

Four scheduling algorithms were studied in this and previous work. The first
is First-Come, First-Served (FCFS) which prioritizes scheduling tasks from
workflows that arrived earlier over tasks from workflows that arrived later.
The second is Proportional-Least Laxity First which projects a finish time
of a task's overarching workflow based on the rate of completion of its tasks
in the past and then prioritizes tasks from workflows projected to be
completed most tardy proportional to the overall size of the workflow. The
final two algorithms are variants of the Cost-Minimization Scheduling Algorithm
(CMSA) \cite{costmin}, which projects a finish time of a task's workflow (as
with PLLF) and uses a cost function to assign a cost to the projected
tardiness. CMSA is used with two cost functions: a quadratic cost function and
sigmoid cost function.

All four studied scheduling algorithms were previously shown to be sensitive to
even small amounts of error in modeled tasks \cite{pdpta18}.  Each scheduling
algorithm exhibited a decreased percent of workflow completed ahead of their
deadline with the smallest amount of error studied.  For three of the
scheduling algorithms the decrease was substantial, but relatively equal,
regardless of the amount of error.  For the fourth scheduling algorithm, the
decrease was less severe overall and was propotionate to the amount of error.
The fourth algorithm was thus declared to be somewhat robust to small amounts
of error (less than 0.5\%) but ultimately, like all three others, was not
robust for errors of 1\% or greater.

The main result of \cite{pdpta18} showed that incorporating feedback of
task completions from the actual platform to the model platform (thereby
preventing the model platform from modeling a task completing before the
actual task completed) dramatically increased robustness of all scheduling
algorithms, even for errors up to 90\% (the highest amount studied) in the
modeled tasks' requirement for amount of work, measured in CPU cycles.  Figure
\ref{fig:underestimating} illustrates why the model platform's underestimation
of the requirements of a task (and thereby modeling it as completed ahead of
the actual task) can be so detrimental.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/UnderestimatedErrorEffect_ModelTaskWork.pdf}
		\includegraphics[width=0.5\textwidth]{figures/UnderestimatedErrorEffect_ActualTaskWork.pdf}
	\end{center}
	\caption{Illustrating the disconnect due to error, assuming model task
		requirements underestimate actual task requirements and the model
		task finishes ahead of the actual task.}
	\label{fig:underestimating}
\end{figure}

In Figure \ref{fig:underestimating} two tasks, $t_1$ and $t_2$, both have
the same amount of CPU cycles, $C$, required for their completion.  The upper
chart illustrates the modeled platform's view of time in which task $t_1$ begins
executing first and when it completed, $t_2$ is scheduled to immediately
begin executing.  However, assuming the modeled requirements of $t_1$ were
an underestimate of the actual $t_1$'s requirements, when the model of
$t_1$ finishes and $t_2$ is scheduled to begin, the actual platform is not
yet finished with $t_1$ and thus must work concurrently on $t_1$ and $t_2$
for some time until the actual $t_1$ task does complete.  This causes both
$t_1$ and $t_2$ to have actual completion times later than the modeled
completion times because of the unintentional over allocation of resources of
the machine executing the tasks.  By extension, if another task, say $t_3$,
were to be scheduled to begin after $t_2$ finishes this problem could compound
because $t_2$'s actual completion would be further delayed by the additional
concurrent execution with $t_3$.

The previous example illustrates why allowing the model platform to
inform scheduling algorithms when tasks are complete and machines are
idle (or less loaded) in the presense of error (which may underestimate tasks'
true requirements) can be so detrimental.  With feedback of every task's
completion from the actual platform, the modeled completion times of tasks
may be safely ignored, in favor of relying on task completion notifications
from the actual platform.  However, complete feedback of all tasks' completion
may be impractical in a live system, or may simply be cost-prohibitive.
This paper thus seeks to address the question of whether partial feedback
of tasks' completion events may be sufficient to achieve some level of
robustness to model error.  In addition, based specifically on the knowledge
of the underestimating problem illustrated prior, Section \ref{sec:Biasing}
proposes a specific approach to counteracting that problem by biasing model
task requirements.  Results of both partial feedback and biasing are presented
in Section \ref{sec:Results}.

\section{Biasing Model Requirements}
\label{sec:Biasing}

As illustrated in Figure \ref{fig:underestimating}, if the model task
requirements are an underestimate of the actual task's requirements then
scheduling algorithms may schedule future tasks to begin executing
unintentionally-concurrent with other tasks, delaying their completions.
However, if the model is known to have error which may underestimate task
requirements, then the model task requirements could simply be biased by
increasing its assumed (provided) value in order to reduce (or ideally,
eliminate) the probability that it underestimates the actual task requirements.
In this section three proposed strategies for biasing task requirements are
proposed.

The simplest form of biasing is to add a constant value to each task
requirement.  If the maximum magnitude of error for which a model
task requirement may underestimate the actual task requirement is known,
then that value would be the ideal bias constant value because it would
eliminate the model from ever underestimating task requirements while
minimizing the amount of overestimation. Practically, the maximum error
magnitude is unlikely to be known. However, because it is the most ideal
circumstance for biasing it is included here and in simulated results of
Section \ref{sec:Results}. The equation for a simple constant bias value, $C$,
addition to each model task requirement, $\hat{X}$, to yield a biased model
task requirement, $\hat{X^b}$, is given in Eq. \ref{eq:cbias}.

\begin{equation}
\hat{X^b} \xleftarrow{c} \hat{X} + C
\label{eq:cbias}
\end{equation}

A more realistic approach to biasing model task requirements is to assume a
bound, not on the magnitude of error, but on the fraction (or percentage) of
error. It may be possible, for example, through analysis of past executions of
tasks and workflows, to estimate the maximum percentage of error for each model
task requirement, $\hat{X}$, relative to the actual task requirement, $X$. In
other words it may be practical to estimate that each model task requirement is
within a factor of $P$ of the actual task requirement.  Although that implies $\hat{X}$
may under- or over-estimate $X$ by as much as $P$, in order to prevent
underestimating we can adjust $\hat{X}$ as in Eq. \ref{eq:pbias}, hereafter
referred to as the proportionate bias strategy.

\begin{equation}
\hat{X^b} \xleftarrow{p} \hat{X} / (1 - P)
\label{eq:pbias}
\end{equation}

Although the proportionate bias strategy can effectively eliminate
underestimated model task requirements with a judiciously chosen $P$
which may require less knowledge about the nature of the error than choosing
a proper magnitude for the constant bias strategy value, $C$, this relaxed
requirement comes at a price. Specifically, where the constant bias
strategy shifts the actual task requirement by $C$, the proportionate bias
strategy `stretches' the distribution of $\hat{X^b}$ and yields a much larger
range of overestimates. This is illustrated in Figure \ref{fig:cpbias} where
the distribution of an example model value is adjusted according to the
constant and proportionate bias strategies.  For both adjustments, an ideal
value of $C$ and $P$ are shown, though practically an ideal value wouldn't be
known and have to be estimated itself.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Constant.pdf}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Proportionate.pdf}
	\end{center}
	\caption{Illustration of probabilities for a model task requirement
		with bound error as a fraction, $P$, of the actual task
		requirement, $X$, and the effect of constant and proportionate
		bias strategies transforming the distribution into
		one that never underestimates the actual term, $X$, given an
		ideal value for $C$ and using the exact value of $P$ (the bounded
		error) in the proportionate bias strategy's Eq. \ref{eq:pbias}.
		Distribution is assumed to be zero-mean and triangular.}
	\label{fig:cpbias}
\end{figure}

In order to achieve a less `stretched' distribution for the biased model task
requirement than the proportionate bias strategy while still maintaining the
need only for an estimated maximum percentage, not magnitude, of error, the
third proposed strategy, known as the simple bias strategy, is given in Eq.
\ref{eq:sbias}. This third strategy fails to eliminate the possibility of
underestimating though it can reduce its propability substantially, but also
prevents wildly overestimating task requirements by decreasing the amount of
`stretch' in the biased term's distribution.  Figure \ref{fig:sbias}
illustrates the effect of the simple bias strategy on the distribution of the
biased term.

\begin{equation}
\hat{X^b} \xleftarrow{s} \hat{X} (1 + P)
\label{eq:sbias}
\end{equation}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Simple.pdf}
	\end{center}
	\caption{Illustration of probabilities for a model task requirement
		with bound error as a fraction, $P$, of the actual task
		requirement, $X$, and the effect of the simple bias strategy
		transforming the distribution, given an ideal value for $P$.
		Distribution is assumed to be zero-mean and triangular.}
	\label{fig:sbias}
\end{figure}

\section{Results}
\label{sec:Results}

All results were collected using simulations performed using simulator software
developed for previous research \cite{pdpta18} and \cite{costmin} and made
publically available as open source \cite{soasim}.  Workflows and task
requirements are the same as those in \cite{pdpta18} as are simulated error
amounts which ranged from $0.1\%$ up to $50\%$ taken from a uniform
distribution.  For all numeric results for performance the value presented is
an averaged value across ten simulations where the workflows and tasks were
identical but an error term applied to model tasks' requirements were unique.

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.9\textwidth,height=3in]{figures/Histogram_All_NoLoss_NoBias_LowAllError.pdf}
	\end{center}
	\caption{Histogram of workflows by normalized tardiness comparing
		relative performance of scheduling algorithms with no error (vertical
		bars) and with 0.1\% error applied to the model of the workflow
		requirements (line graphs).  Complete feedback of task completion from
		the actual platform was used, but model platform was, due to
		underestimated task requirements, allowed to model tasks as completing
		and, as a result, scheduling algorithms were allowed to schedule
		additional tasks to the machine.}
	\label{fig:noloss-nobias-lowerror}
\end{figure*}

Figure \ref{fig:noloss-nobias-lowerror} shows the performance of the four
scheduling algorithms and the significant performance impact that the smallest
amount of error tested has even when complete feedback is available from the
actual platform but the model platform is still allowed to model tasks as
completing early due to underestimates of task requirements. (In other words,
the model platform only utilizes actual completion events when they occur
before modeled completion events.)  In this figure
as with prior research, performance is depicted visually as a histogram of the
number of workflows completed in intervals based on their normalized tardiness
(the time difference between the completion and the target deadline normalized
by amount of time available to execute the workflow, i.e., the difference of
deadline and arrival time of the workflow). In this representation a normalized
tardiness of 0 represents a workflow that completed exactly at its deadline,
negative values represent workflows completed before their deadline, and
positive values those completed late.

The histogram bars of Figure \ref{fig:noloss-nobias-lowerror} represent the
performance of scheduling algorithms under the ideal circumstance of no error
in the model (i.e., the model platform perfectly predicts and represents the
resources required by a task and its execution completion time).  These
histogram bar results demonstrate how CMSA with either of the two
cost functions (Sigmoid or Quadratic) completes the largest majorities of
workflows ahead of their deadline.  The PLLF (proportional least laxity first)
algorithm completes workflows up to 4 times later (as a proportion of the
ideal finish time) than the deadline. The FCFS (first-come, first-serve)
algorithm performs relatively poorly with workflows completing far later than
their deadline because the algorithm doesn't use deadline information in making
its scheduling decisions.

The lines graphed in Figure \ref{fig:noloss-nobias-lowerror} represent the
same algorithms' histogram of workflow completion in the presence of 0.1\% error
in the model platform.  Due even to this small amount of error, the problem of
underestimating task requirements and overallocating resources, and the nature
of this issue compounding results in all four scheduling algorithms exhibiting
substantially worse performance.  This is illustrated by the reduction to less
than half as many workflow completed before their deadlines (normalized
tardinesses less than 0) compared to the no-error case.  It is also
demonstrated by the large increase of number of workflows completed with a
normalized tardiness of 10 or higher compared to the no-error case.

Therefore, with any level of partial feedback available none of these
scheduling algorithms would perform better than the case of
complete feedback being available but still allowing modeling of early task
completions due to underestimated task requirements.  However, the use of
model task requirement biasing showed promising results at restoring
performance of the algorithms by eliminating or reducing the likelihood of
the underestimating problem.

Figure \ref{fig:nofeedback-cbias-higherror} depicts the effect of the constant
bias strategy in the present of the highest ($50\%$) error tested with no
feedback (the lines of the graph) against the no-error ideal case (the bars
of the graph).  Key results are that for constant bias strategy is of no
help to the FCFS algorithm which completes the majority of workflows with a
normalized tardiness of $10$ or above.  For the PLLF algorithm, nearly as
many workflows are no longer completed on time as for FCFS but the maximum
normalized tardiness of workflows remains bounded at about $4$.  Both
variants of CMSA seem to similarly complete fewer overall workflows on time
through for the Quadratic cost function variant those workflows seem to be
completed mostly at $0.5$ normalized tardiness and below while for the
Sigmoid cost function variant they are spread from normalized tardinesses
under $0.25$ up to $10+$. Although not depicted, for smaller bounded errors,
the constant bias strategy does expectedly better, achieving results nearly
as good as the no-error ideal case for errors up to about $5\%$.

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.9\textwidth,height=3in]{figures/Histogram_All_CompleteHighAllUniformError_WithCBias.pdf}
	\end{center}
	\caption{Histogram of workflows by normalized tardiness comparing
		relative performance of scheduling algorithms in ideal case with no
		error (vertical bars) and with 50\% error applied to the model of the
		workflow requirements (line graphs) given no feedback of task
		completion times and using the constant bias strategy with ideal
		constant to prevent underestimating task requirements in even the
		worst instance of error.}
	\label{fig:nofeedback-cbias-higherror}
\end{figure*}

Neither the proportionate nor simple bias strategies achieve acceptible
results with errors as high as $50\%$.  In Figures
\ref{fig:nofeedback-pbias-mederror} and \ref{fig:nofeedback-sbias-mederror}
the effect of proportionate and simple bias strategies are shown, respectively,
in the case of $5\%$ error with no feedback and ideal bias value $P=5\%$.
For the proportionate bias strategy, all four scheduling algorithms complete
about one third fewer workflows at a normalized tardiness of $-0.25$ and below
although most of those workflows are completed at normalized tardiness up to
zero, which is still before their deadline.  CMSA with both cost functions completes
a few more workflows mildly late (normalized tardiness $0.25$ and below), PLLF
completes more workflows at normalized tardinesses up to $1$, and FCFS a few
more at $4+$.  Overall, the proportionate bias strategy fairly effectively
makes all scheduling algorithms robust to errors up to $5\%$ even without any
actual platform feedback available.

The simple bias strategy achieves an even better result than the proportionate
bias strategy for the $5\%$ bounded error case, showing fewer workflows
shifting from being completed early to completed late.  In fact, the CMSA with
Sigmoid cost function appears to achieve nearly identical results despite no
feedback and up to $5\%$ error when using the simple bias strategy.  It
therefore appears that the small chance for simple bias strategy to still
underestimate task requirements is a better tradeoff than the exagerrated
overestimates produced by the proportionate bias strategy.

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.9\textwidth,height=3in]{figures/Histogram_All_CompleteMediumAllUniformError_WithRBias.pdf}
	\end{center}
	\caption{Histogram of workflows by normalized tardiness comparing
		relative performance of scheduling algorithms in ideal case with no
		error (vertical bars) and with 5\% error applied to the model of the
		workflow requirements (line graphs) given no feedback of task
		completion times and using the proportionate bias strategy with ideal
		$P=5\%$ to prevent underestimating task requirements in even the
		worst instance of error.}
	\label{fig:nofeedback-pbias-mederror}
\end{figure*}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.9\textwidth,height=3in]{figures/Histogram_All_CompleteMediumAllUniformError_WithBias.pdf}
	\end{center}
	\caption{Histogram of workflows by normalized tardiness comparing
		relative performance of scheduling algorithms in ideal case with no
		error (vertical bars) and with 5\% error applied to the model of the
		workflow requirements (line graphs) given no feedback of task
		completion times and using the simple bias strategy with ideal
		$P=5\%$ to nearly prevent underestimating task requirements in even the
		worst instance of error.}
	\label{fig:nofeedback-sbias-mederror}
\end{figure*}

\section{Summary and Future Work}
\label{sec:Summary}

The case for why complete and reliable feedback of scheduled task completions
from an actual platform to the model used for scheduling decision-making was
found to be necessary to achieve robust performance of scheduling algorithms,
as first described in previous work \cite{pdpta18}. Furthermore we have
demonstrated the use of partial feedback to be insufficient at achieving
similar robustness, but have proposed the use of biasing the model parameters
to prevent or reduce the likelihood of the model underestimating task
requirements thereby modeling tasks as completing earlier than their actual
completion. Three biasing strategies were proposed with unique trade-offs
of assumed knowledge about error bounds and degree to which they decrease
underestimation and increase overestimation of task requirements. Through
simulated case studies we demonstrated that each biasing strategy is useful
for achieving robustness of scheduling algorithm performance at different
amounts of model error.

Future work may include combination of partial feedback with biasing
strategies.  Additionally, other biasing strategies or alternatives to
biasing will be explored to increase robustness or address cases where
the amount of error in the model is unknown.

\bibliographystyle{IEEEtran}
\bibliography{paper}{}

\end{document}